---
title: "Cruising speed versus average speed"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE)
```


## Required packages

Packages that needs to be installed:

```{r}
from_cran <- c("dplyr", "parallel", "purrr", "tidyr", "readr", "R.utils", "stringr")
to_install <- setdiff(from_cran, installed.packages()[, "Package"])
if (length(to_install)) install.packages(to_install)
```

The `FITfileR` package also needs to be installed from
[Github](https://github.com/grimbough/FITfileR). Loading some of these packages:

```{r}
library(FITfileR)
library(dplyr)
library(purrr)
library(parallel)
```


## Downloading Strava activities

* Hover over your name in the upper right-hand corner of the Strava page. Choose
`Settings`, then find the `My Account` tab from the menu listed on the Left.

* Select `Get Started` under `Download or Delete Your Account`

* Select `Request your archive` on the next page.

You will receive an email with a link to download your data (this may take a few
hours). For this reason, it’s important that you have access to the email
account attached to your Strava profile. More info
[here](https://support.strava.com/hc/en-us/articles/216918437-Exporting-your-Data-and-Bulk-Export#h_01GG58HC4F1BGQ9PQZZVANN6WF).

I've put my Strava data here:

```{r}
data_folder <- "~/Library/CloudStorage/OneDrive-OxfordUniversityClinicalResearchUnit/data/Strava/"
```

and the downloaded data are in this subfolder:

```{r}
raw_strava <- "export_34225766/"
```

Retrieving the names of the files that correspond to cycling activities since
2022:

```{r}
# files <- "~/Desktop/export_34225766/activities.csv" |>
files <- paste0(data_folder, raw_strava, "activities.csv") |>
  readr::read_csv() |> 
  tidyr::separate_wider_delim(`Activity Date`, ",",
                              names = c("day", "year", "time")) |> 
  mutate(across("year", as.integer)) |> 
  filter(`Activity Type` == "Ride", year > 2021) |> 
  pull(Filename)
```

Keeping only the FIT files:

```{r}
files <- files[grep("fit", files)]
```

Let's copy the cycling activities to another folder in which we will decompress
them:

```{r eval = FALSE}
riding_folder <- paste0(data_folder, "activities/")
dir.create(riding_folder)
dest <- paste0(riding_folder, files)
file.copy(paste0(data_folder, raw_strava, files), dest)
walk(dest, R.utils::gunzip)
```


## Some utilitary functions

Converts speed from m/s to km/h:

```{r}
ms2kmh <- function(x) 3.6 * x
```

Extracts the speeds values from the data structure:

```{r}
extract_speeds <- function(x) {
  idx <- grep("speed", names(x))
  if (length(idx)) return(as.vector(x[[idx]]))
  NULL
}
```

Makes sure that the output of `records()` is a list:

```{r}
make_list <- function(x) {
  if (is.data.frame(x)) return(list(x))
  x
}
```

Retrieves speeds from an activity:

```{r}
get_speed <- function(x) {
  x |> 
    readFitFile() |> 
    records() |> 
    make_list() |> 
    map(extract_speeds) %>%
    do.call(c, .) |>  
    ms2kmh()
}
```

Computes the cruising speed as the mode of the distribution of speeds:

```{r}
cruising_speed <- function(x) {
  dens <- density(x)
  dens$x[which.max(dens$y)]
}
```


## Example on one activity

The blue curve is the distribution of speeds during the activity and the red and
green vertical lines show the average and cruising speeds respectively:

```{r show1activity}
sp <- get_speed(paste0(data_folder, "activities/10517736664.fit"))
dens <- density(sp)
plot(dens, xlim = c(0, max(dens$x)), ylim = c(0, 1.03 * max(dens$y)),
     xaxs = "i", yaxs = "i",
     main = NA, xlab = "speed (km/h)", ylab = "density", col = 4, lwd = 3)
(asp <- mean(sp))
(csp <- cruising_speed(sp))
abline(v = c(asp, csp), col = 2:3, lwd = 2)
```

It shows for example that the average speed is `r round(asp, 1)` km/h (red
vertical line) whereas the cruising speed is actually `r round(csp, 1)` km/h
(green vertical line).


## Processing all the cycling acitivities

Computes the average and cruising speeds from a vector of speeds:

```{r}
speeds <- function(x) {
  c(mean, cruising_speed) |> 
    map_dfc(function(f) f(x)) |> 
    setNames(c("average", "cruising"))
}
```

Reads a compressed FIT file from disk and compute the average and cruising
speeds from it:

```{r}
read_file <- function(x) {
  x |> 
    get_speed() |> 
    speeds()
}
```

Processing all the files:

```{r eval = FALSE}
speeds_lst <- files |> 
  stringr::str_remove(".gz") %>%
  paste0(data_folder, .) |> 
  mclapply(read_file, mc.cores = detectCores() - 1) |> 
  bind_rows()
```

```{r include = FALSE}
speeds_lst <- readRDS(paste0(data_folder, "speeds_lst.rds"))
```

Modelling the cruising speed as a function of the average speed:

```{r}
mod <- lm(cruising ~ average, speeds_lst)
```

The model's parameters estimates are the following:

```{r}
coef(mod)
```

Which means that, as a rule of thumb, you can obtain cruising speed by adding
6 km/h to 90% of the value of the average speed. But, more precisely, the
following function converts the cruising speed into average speed from the above
model:

```{r}
intercept <- coef(mod)[1]
slope <- coef(mod)[2]

cs2av <- function(x) {
  (x - intercept) / slope
}
```

The cruising speeds that define the groups:

```{r}
crsp <- c(40, 37, 41, 34, 38, 31, 35, 25, 30)
```

A function that nicely formats speed ranges:

```{r}
speed_range <- function(x) {
  c(paste0(x[1], "+ km/h"),
    map_chr(as.data.frame(matrix(x[-1], 2)),
            function(y) paste0(y[1], "-", y[2], " km/h")))
}
```

The comparison of cruising and average speeds for the groups:

```{r}
avsp <- cs2av(crsp)

data.frame(
  group          = c("A", "B+", "B", "C", "D"),
  cruising_speed = speed_range(crsp),
  average_speed  = speed_range(round(avsp))
)
```

Let's see the data points and the model:

```{r}
opar <- par(pty = "s")
axlim <- c(18, 42)
nb <- length(crsp)

with(speeds_lst,
     plot(average, cruising, col = 4, xlim = axlim, ylim = axlim,
          xlab = "average speed (km/h)", ylab = "cruising speed (km/h)"))
abline(mod, col = 2, lwd = 2)

segments(rep(0, nb), crsp, avsp, crsp)
segments(avsp, rep(0, nb), avsp, crsp)
box(bty = "o")

par(opar)
```


