---
title: "Importing Strava data"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preamble

Needed packages:

```{r message = FALSE}
library(stringr)
library(FITfileR)
library(dplyr)
library(magrittr)
```

The Strava data are here:

```{r}
data_folder <- "~/Library/CloudStorage/OneDrive-OxfordUniversityClinicalResearchUnit/data/Strava/"
```

```{r eval = FALSE, include = FALSE}
exportations <- data.frame(date = lubridate::ymd(20230912),
                           name = "export_34225766/")

saveRDS(exportations, paste0(data_folder, "exportations.rds"))
```

and the downloaded raw data are in this subfolder:

```{r}
raw_strava <- data_folder |>
  paste0("exportations.rds") |>
  readRDS() |> 
  arrange(date) |> 
  pull(name) |> 
  last()
```

## Activities summary

Let's load the summary table of the of activities:

```{r}
activities <- readr::read_csv(paste0(data_folder, raw_strava, "activities.csv"))
```

Fixing some problems regarding some names:

```{r}
tmp <- names(activities)
idx <- grep("horton", tmp)
foo <- tmp[idx]
bar <- foo |> 
  str_remove("^.*\">") |> 
  str_remove("<.*$")

if (length(intersect(bar, tmp)) < 1) {
  tmp[idx] <- bar
  names(activities) <- tmp
}
```

Some columns are more or less duplicated. Let's remove them by choosing the one
with most information:

```{r}
to_discard <- function(x) {
  subact <- select(activities, contains(x))
  the_names <- names(subact)
  
  check <- subact |> 
    setNames(c("a", "b")) |> 
    mutate(diff = a - b) |> 
    filter(diff != 0 | is.na(diff))
  
  tmp <- apply(check, 2, function(x) sum(is.na(x)))
  rslt <- which(tmp < 1)
  lgth <- length(rslt)
  
  if (lgth) {
    if (lgth < 2) {
      return(the_names[-rslt])
    }
  } else {
    return(the_names[which.max(tmp[-3])])
  }
}

to_remove <- grep("\\d", names(activities), value = TRUE) |> 
  str_remove("\\d+") |> 
  unique() |> 
  purrr::map_chr(to_discard)

activities[to_remove] <- NULL
names(activities) <- str_remove(names(activities), "...\\d+$")
```

Final tuning and cleaning:

```{r}
activities <- activities |> 
  mutate_at("Activity Date", lubridate::mdy_hms) |> 
  arrange(`Activity Date`) |> 
  select(-`Prefer Perceived Exertion`, -`From Upload`) |> 
  select(where(~ !all(is.na(.x))))
```

Note here that the date and time is in UTC. Note also that not all of them are
in FIT format:

```{r}
activities |> 
  pull(Filename) |> 
  str_remove(".gz") |> 
  str_remove("^.*\\d.") |> 
  table()
```

## FIT files

Let's start by reading the FIT files first. Here is the names of the FIT files:

```{r}
files <- activities |> 
  pull(Filename) %>% 
  grep("fit", ., value = TRUE)
```

Next, we copy all the FIT files to another directory where we unzip them:

```{r eval = FALSE}
destination <- paste0(data_folder, str_replace(raw_strava, "export", "fit"))
dir.create(destination)
dest <- paste0(destination, str_remove(files, "activities/"))
file.copy(paste0(data_folder, raw_strava, files), dest)
walk(dest, R.utils::gunzip)
```

```{r include = FALSE}
destination <- paste0(data_folder, str_replace(raw_strava, "export", "fit"))
```

Let's now read each of the unzipped FIT file and convert them into an RDS file
that we save into another folder:

```{r eval = FALSE}
rds_files <- str_replace(destination, "fit", "rds")
dir.create(rds_files)

files <- files |> 
  str_remove("activities/") |> 
  str_remove(".fit.gz")
  
fit2rds <- function(x) {
  tmp <- purrr::safely(readFitFile)(paste0(destination, x, ".fit"))
  if (is.null(tmp$error)) {
    saveRDS(tmp$result, paste0(rds_files, x, ".rds"))
  }
  rm(tmp)
}

walk(files, fit2rds)
```

```{r include = FALSE}
rds_files <- str_replace(destination, "fit", "rds")
```

Errors occurred for a number of files:

```{r}
error_files <- setdiff(str_remove(dir(destination), ".fit"),
                       str_remove(dir(rds_files), ".rds"))
length(error_files)
head(error_files)
```

Just confirming that it doesn't work:

```{r eval = FALSE}
readFitFile(paste0(destination, "2356152777", ".fit"))
```

Not even after having unzipped the file by hand:

```{r eval = FALSE}
readFitFile("~/Desktop/2356152777.fit")
```

One possibility is that there has been errors either when the archive was
created by Strava, or during the downloading process. For the latter we can just
try downloading the archive again. For the former we can make another request in
one week and download the new archive.

## GPX and TCX files

```{r}
gpx_dates <- activities |>
  filter(grepl("gpx", Filename)) |>
  pull(`Activity Date`)

nearby_activities <- function(x) {
  filter(activities, `Activity Date` > as_date(x) - 2, `Activity Date` < as_date(x) + 2)
}

(a <- map(gpx_dates, nearby_activities))
```

There are only activities `1946738728` and `2211797944` that we can suspect to
be identical, as well as `2211797355` and `1950384218`, even though the names of
the files are different:

```{r}
pull(a[[1]], Filename)
```

Let's see with TCX:

```{r}
tcx_dates <- activities |>
  filter(grepl("tcx", Filename)) |>
  pull(`Activity Date`)

map(tcx_dates, nearby_activities)
```

Something weird between 2020-07-17 and 2020-07-28

```{r eval = FALSE}
library(activatr)

gpxfile <- parse_gpx("~/Desktop/1946738728.gpx")
mutate_with_speed(gpxfile)
```

